---
title: Primer Dataset Ideas
author: Ezgi Karaesmen
date: '2020-05-27'
slug: primer-ideas
output: 
  blogdown::html_page:
    toc: true
categories: [primer]
tags:
  - interactive primers
  - datasets
---



# 1. Build a Model

## Good practice

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(funModeling)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, error=FALSE)
```


### 1. Outcome: numeric

Sad truth about professor salaries from `carData`. [See doc](https://vincentarelbundock.github.io/Rdatasets/doc/carData/Salaries.html)

```{r}
salaries <- read_csv("https://vincentarelbundock.github.io/Rdatasets/csv/carData/Salaries.csv")
glimpse(salaries)
```


### 1. Outcome: binary

Esophagial cancer

```{r}
glimpse(esoph)
```

either expand the dataset to reflect individual observations or work with case/control probabilities?
e.g. a tidymodels version of

```{r, eval=FALSE}
model1 <- glm(cbind(ncases, ncontrols) ~ agegp + tobgp * alcgp,
              data = esoph, family = binomial())
anova(model1)
```


### 1. Outcome: categorical

Is there a dataset with SAT scores and career choices?

Some other suggestions:

Canadian Women's Labour-Force Participation [See doc.](https://vincentarelbundock.github.io/Rdatasets/doc/carData/Womenlf.html)

```{r}
wforce <- read_csv("https://vincentarelbundock.github.io/Rdatasets/csv/carData/Womenlf.csv")
glimpse(wforce)
wforce %>% 
  group_by(partic, children) %>%
  count()
```


Pottery: Chemical composition of ancient pottery found at four sites in Great Britain. [See doc.]

```{r}
pott <- read_csv("https://vincentarelbundock.github.io/Rdatasets/csv/carData/Pottery.csv")
glimpse(pott)
```



# 1. Preprocessing

## Good practice

### Working with PCA

```{r}
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')
table(spotify_songs$playlist_genre)
hist(spotify_songs$track_popularity)
glimpse(spotify_songs)
```





