---
title: Preprocessing-GP Continued...
author: Ezgi
date: '2020-05-28'
slug: preprocessing-gp-continued
output: 
  blogdown::html_page:
    toc: true
categories:
  - primer
tags:
  - preprocessing
  - good practice
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE, warning = FALSE)
library(tidyverse)
library(tidymodels)
```


Let's keep checking out some other dataset shall we?

# Pottery dataset

The data give the chemical composition of ancient pottery found at four sites in Great Britain. [See doc](https://vincentarelbundock.github.io/Rdatasets/doc/carData/Pottery.html)

```{r}
pottery <- read_csv("https://vincentarelbundock.github.io/Rdatasets/csv/carData/Pottery.csv")
glimpse(pottery)
summary(pottery)

pottery %>%
  count(Site)


pca_rec <- recipe(~., data = pottery) %>%
  update_role(X1, Site, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

pca_prep <- prep(pca_rec)
```

## Contribution by variable

```{r}
tidied_pca <- tidy(pca_prep, 2)

tidied_pca %>%
  filter(component %in% paste0("PC", 1:5)) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = terms)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component, nrow = 1) +
  labs(y = NULL)

library(tidytext)

tidied_pca %>%
  filter(component %in% paste0("PC", 1:4)) %>%
  group_by(component) %>%
  top_n(8, abs(value)) %>%
  ungroup() %>%
  mutate(terms = reorder_within(terms, abs(value), component)) %>%
  ggplot(aes(abs(value), terms, fill = value > 0)) +
  geom_col() +
  facet_wrap(~component, scales = "free_y") +
  scale_y_reordered() +
  labs(
    x = "Absolute value of contribution",
    y = NULL, fill = "Positive?"
  )

```

## PCA

```{r}
juice(pca_prep) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = Site), alpha = 0.7, size = 2)
```

## UMAP

```{r}
set.seed(9876)

library(embed)
umap_rec <- recipe(~., data = pottery) %>%
  update_role(X1, Site, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_umap(all_predictors())

umap_prep <- prep(umap_rec)

juice(umap_prep) %>%
  ggplot(aes(umap_1, umap_2)) +
  geom_point(aes(color = Site), alpha = 0.7, size = 2)

```


## Actual Map

If iframe is giving me a hard time here's the [map link](https://www.google.com/maps/d/drive?state=%7B%22ids%22%3A%5B%221LpiNnGfzSmi8lkOsFDvGbJkf7JMjfvAu%22%5D%2C%22action%22%3A%22open%22%2C%22userId%22%3A%22116929632846172529102%22%7D&usp=sharing)

<iframe src="https://www.google.com/maps/d/u/0/embed?mid=1LpiNnGfzSmi8lkOsFDvGbJkf7JMjfvAu" width="640" height="480"></iframe>

# Wine quality

The Wine Quality Dataset involves predicting the quality of white wines on a scale given chemical measures of each wine. [See](http://archive.ics.uci.edu/ml/datasets/Wine+Quality)



```{r}
foo1 <- function(x) ifelse(str_detect(x, "^[.]"), paste0("0", x), x) 
foo2 <- function(x) map_chr(x, foo1)

wine <- read_csv2("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv")

wine %>%
  mutate_if(is.character, foo2) %>%
  mutate_if(is.character, as.numeric) -> wine

glimpse(wine)
summary(wine)

pca_rec <- recipe(~., data = wine) %>%
  update_role(quality, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

pca_prep <- prep(pca_rec)
```

## Contribution by variable

```{r}
tidied_pca <- tidy(pca_prep, 2)

tidied_pca %>%
  filter(component %in% paste0("PC", 1:5)) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = terms)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component, nrow = 1) +
  labs(y = NULL)

library(tidytext)

tidied_pca %>%
  filter(component %in% paste0("PC", 1:4)) %>%
  group_by(component) %>%
  top_n(8, abs(value)) %>%
  ungroup() %>%
  mutate(terms = reorder_within(terms, abs(value), component)) %>%
  ggplot(aes(abs(value), terms, fill = value > 0)) +
  geom_col() +
  facet_wrap(~component, scales = "free_y") +
  scale_y_reordered() +
  labs(
    x = "Absolute value of contribution",
    y = NULL, fill = "Positive?"
  )
```

## PCA

```{r}
juice(pca_prep) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = quality), alpha = 0.7, size = 2) +
  scale_color_viridis_c()
```

## UMAP

```{r}
library(embed)

set.seed(9876)

umap_rec <- recipe(~., data = wine) %>%
  update_role(quality, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_umap(all_predictors())

umap_prep <- prep(umap_rec)

juice(umap_prep) %>%
  ggplot(aes(umap_1, umap_2)) +
  geom_point(aes(color = quality), alpha = 0.7, size = 2) +
  scale_color_viridis_c()
```

